# PA1. Reproducible Research, by Coursera
# amaurandi@um.es
# 5.Junio.2014
----------------------------------------
```{r setopts, echo =FALSE}
opts_chunk$set(echo=T, results="asis")
options(scipen = 1, digits = 3)  # set default digits
library(xtable)

```


# Loading and processing the data

Exploration of the data and data structure.  
  
```{r readdata, echo=T, results="asis"}
df<-read.table("activity.csv", sep=",", header=T)
sdf<-summary(df)
xt<-xtable(sdf)

print(xt, type="html")

```

He have arround 288 observations per day.

# What is mean total number of steps taken per day?


```{r desc, echo=T}
# no account of misssign values
df1<-na.omit(df)
su<-tapply(df1$steps, df1$date, sum)
su<-as.numeric(su)
```

Histogram of total number of steps per day
```{r histograma, fig.align='center',fig.height=5, fig.width=5}
hist(su, main="Histogram of the total steps by day", breaks = 50, freq=T)
```


```{r descrip, echo=FALSE}
# mean and median
m<-mean(su, na.rm=T)
m2<-median(su,  na.rm=T)

```

The mean of the total step per day is `r m` and the median is `r m2 `.

# What is the average daily activity pattern?










```{r}
# save a numeric vector containing 48 monthly observations
# from Jan 2009 to Dec 2014 as a time series object
myts <- ts(myvector, start=c(2009, 1), end=c(2014, 12), frequency=12) 

# subset the time series (June 2014 to December 2014)
myts2 <- window(myts, start=c(2014, 6), end=c(2014, 12)) 

# plot series
plot(myts)
```












# Imputing missing values


# Are there differences in activity patterns between weekdays and weekends?











```{r}
sessionInfo()
```

















